Virtual Memory Assisted Transactional Memory for Dynamic Languages

### 2014-05-27 talk/dls2014/paper/paper.tex
remove rss comparison and try with some page privatisation statistics

Third, since we do copy on write, if all objects were to be
written to at the same time, all pages would need to be privatised
for all objects in the old object space. In that case we would
need the total amount of memory required by old objects multiplied
by $N+1$ (incl. the sharing segment). Pages get re-shared during
major collections if possible.
\remi{maybe collect some statistics about pages privatised per segment}

\remi{The following discussion about richards mem usage does not
say that much... Also, RSS is not a good measure but it's hard to
get something better.}
In figure \ref{fig:richards_mem} we look at the memory usage of
one of our benchmarks called Richards\footnote{OS kernel simulation
benchmark}. The \emph{Resident Set Size} (RSS) shows the physical memory
assigned to the process. From it, we see that the process' memory
usage does not explode during the benchmark but actually stays pretty
much the same after start-up. Since it is the job of the OS to map
physical memory, this RSS number should be seen as a maximum. It is
possible that some of the memory is not required any more but still
assigned to our process.

The \emph{GC managed memory} counts all memory used in the old object
space including the memory required for private pages. The sharp drops
in memory usage come from major collections that free old objects and
re-share pages. Again the overall memory usage stays the same and
we see that in this benchmark we have around 1 major collection every
second.

For PyPy-STM the average memory requirement is 29~MiB and there are
$\sim 11$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time.

We are missing a memory optimisation to store small objects in a more
compact way, which is done by a normal PyPy not using STM.
Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{plots/richards_mem.pdf}
  \caption{Actual memory managed by the GC and resident set size
    over time in Richards benchmark\label{fig:richards_mem}}
\end{figure}

@@

\remi{gs:segment prefix overhead is virtually none (maybe instruction cache)}
\remi{update numbers in pypy/TODO}

>>>>

Third, since we do copy on write, if all objects were to be
written to at the same time, all pages would need to be privatised
for all objects in the old object space. In that case we would
need the total amount of memory required by old objects multiplied
by $N+1$ (incl. the sharing segment). Pages get re-shared during
major collections if possible.

In figure \ref{fig:richards_mem} we look at the memory usage of one of
our benchmarks called Richards\footnote{OS kernel simulation
benchmark}. The \emph{GC managed memory} counts all memory used in the
old object space including the memory required for private pages. The
sharp drops in memory usage come from major collections that free old
objects and re-share pages. The average memory usage stays around
29~MiB and we see that in this benchmark we have around 1 major
collection every second. The \emph{page privatisation}, which
represents the percentage of used pages with at least one private
copy, shows the same spikes as the memory usage. These come directly from
re-sharing the pages. The maximum page privatisation is around $20\%$
between major collections. Thus we can say that $~20\%$ of the old
objects get modified between collections in this benchmark.

For PyPy-STM the average memory requirement is 29~MiB and there are
$\sim 11$ major collections during the runtime. Normal PyPy with a GIL
grows its memory up to just 7~MiB and does not do a single major
collection in that time. Compared to normal PyPy, we are missing a
memory optimisation to store small objects in a more compact
way. Additionally, since normal PyPy uses a GIL, it does not need to
duplicate any data structures like e.g. the Nursery for each
thread. This, the missing optimisation, and the additional memory
requirements for STM explained above account for this difference.
\remi{I don't know how much sense it makes to go deeper. We will
improve this in the future, but right now this is the overall picture.}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\columnwidth]{plots/richards_mem.pdf}
  \caption{Actual memory managed by the GC and the page privatisation
    over time in Richards benchmark\label{fig:richards_mem}}
\end{figure}

@@

\remi{do it on a non-jit build (see reason above)}
\remi{gs:segment prefix overhead is virtually none (maybe instruction cache)}
\remi{update numbers in pypy/TODO}
